---
title: "Number of Admissions in 2018 (Poisson Model)"
output: 
  html_document:
    toc: true
    toc_float: true
    
---

```{r setup, include = FALSE}
## load packages
library(data.table)
library(ggplot2)
library(drake)
library(kableExtra)
library(countreg)
library(skimr)
devtools::load_all("../../../savvyR")

## loadd processesd files
loadd(df_mdl)
loadd(plt_model_coefs)
loadd(rmse_df)
loadd(conf_mt_cutoff0_test)
loadd(conf_mt_cutoff0_train)
loadd(conf_mt_cutoff1)
loadd(conf_mt_cutoff2)

model_names <- names(plt_model_coefs)


knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Function to be used for dynamic tabsetting
catHeader = function(text = "", level = 3) {
  cat(paste0("\n\n", 
             paste(rep("#", level), collapse = ""), 
             " ", text, "\n"))
}

```



## Population of Interest Characteristics

```{r, fig.height = 4, fig,width = 6}
ggplot(df_mdl, aes(admit_counts)) +
  geom_histogram(bins = 50) + 
  theme_rnd()
```

```{r, eval = FALSE}
df_mdl[,eq_0 := ifelse(admit_counts == 0, 1, 0)]
df_mdl[,.N,by=.(age_bin, eq_0)][order(age_bin)]
```

```{r, eval=FALSE}
loadd(frail_cols)
lapply(df_mdl[,..frail_cols], table)
```

```{r}
df_mdl[,.(age_bin_ = cut(age_2017, seq(65, 110, by = 5), include.lowest = TRUE))] %>%  
  .[,.N, age_bin_] %>% 
  .[,pct := N/length(unique(df_mdl$savvy_pid)) * 100] %>% 
  .[] %>% 
  kable() %>% 
  kable_styling(full_width = FALSE)
  
  

```

```{r}
df_mdl[,.N, gender][,pct := N/length(unique(df_mdl$savvy_pid))][]

```

```{r}
loadd(frail_cols)
melt(df_mdl[,..frail_cols][,lapply(.SD, sum)])[,pct := value/length(unique(df_mdl$savvy_pid)) * 100][] %>% 
  kable() %>% 
  kable_styling(full_width = FALSE)
```

```{r}
loadd(cc_cols)
melt(df_mdl[,..cc_cols][,lapply(.SD, sum)])[,pct := value/length(unique(df_mdl$savvy_pid)) * 100][order(-value)] %>% 
  .[1:10] %>% 
  kable() %>% 
  kable_styling(full_width = FALSE)
```


<br>


## Model Coefficients
### Poisson Model (No Overdispersion) {.tabset}
```{r, results = 'asis', echo = F, fig.width = 10, fig.height = 8}
idx <- grep("poisson", names(plt_model_coefs))

for (i in 1:length(idx)) 
{
  catHeader(model_names[idx[i]], 4)
  print(plt_model_coefs[[idx[i]]]$count)
}
```

### Poisson Model (With Overdispersion) {.tabset}

```{r fig.width = 10, fig.height = 8, results = 'asis'}

idx <- grep("pois w/ disp", names(plt_model_coefs))
for (i in 1:length(idx)) 
{
  catHeader(model_names[idx[i]], 4)
  print(plt_model_coefs[[idx[i]]]$count)
}

```

### Zero-inlated models (count model) {.tabset}

```{r, fig.width = 10, fig.height = 8, results = 'asis'}

idx <- grep("zero inf|zif", names(plt_model_coefs))
for (i in 1:length(idx)) 
{
  catHeader(model_names[idx[i]], 4)
  print(plt_model_coefs[[idx[i]]]$count)
}

```

## Model Checking

### Rootograms {.tabset}
```{r, fig.width = 8, fig.height = 4, results = 'asis', eval = FALSE}

for (i in 1:length(model_list_all))
{
   
  fm <- model_list_all[[i]]$family$family
  nl <- is.null(fm)
  cl <- class(model_list_all[[i]])
  
  if("zeroinfl" %in% cl)
  {
    catHeader(model_names[i], level = 4)
    rootogram(model_list_all[[i]], main = "")
    
  }
  
  if (!nl)
  {
     if (fm != 'quasipoisson')
    {
      catHeader(model_names[i], level = 4)
      rootogram(model_list_all[[i]], main ="")
      
    }
    
  }
  
}


```



## Model Performance Evaluation {.tabset}

### Root-mean squared error
```{r, fig.width = 6, fig.height = 4}
setorderv(rmse_df, 'RMSE')
rmse_df[,RMSE := round(RMSE, 3)]
rmse_df %>% 
kable() %>% 
  kable_styling(full_width = F)
```

<br>

### % Precision and Recall {.tabset}

Positive classes are based on > count threshold predictions, and are all based on test data.

```{r, echo = F}
get_stats = function(cf_list, attr, value_name)
{
melt(as.data.table(data.frame(
  lapply(cf_list, function(x) x$byClass[attr]),
  check.names = FALSE
  )),
  value.name = value_name, 
  variable.name = "model_type")
  
}


pr_recall_kable = function(cf_list)
{
  ppv_df <- get_stats(cf_list, "Pos Pred Value", "ppv")
  recall_df <- get_stats(cf_list, "Recall", "recall")
  setorder(ppv_df, -ppv)
  setorder(recall_df, -recall)
  
  stats_df <- merge(ppv_df, recall_df, by = "model_type")
  stats_df[,c("ppv", "recall") :=  lapply(.SD, function(x) round(x * 100, 2)), 
           .SDcols = c("ppv", "recall")]
  stats_df %>% 
    kable() %>% 
    kable_styling(full_width = F)
}


```
#### > 0 counts

```{r}
pr_recall_kable(conf_mt_cutoff0_test)

```

#### > 1 counts

```{r}
pr_recall_kable(conf_mt_cutoff1)

```

#### > 2 counts
```{r}

pr_recall_kable(conf_mt_cutoff2)
```

### Positive Predictions {.tabset}

#### Train (> 0 counts)

```{r}
melt(as.data.table(lapply(conf_mt_cutoff0_train, function(x)
  round(sum(x$table[1, ]) / sum(x$table) * 100, 3))), variable.name = "model_type", 
  value.name = "% Positive Predictions") %>% 
  kable() %>% 
  kable_styling(full_width = F)
```


#### Test ( > 0 counts)

```{r}
melt(as.data.table(lapply(conf_mt_cutoff0_test, function(x)
  round(sum(x$table[1, ]) / sum(x$table) * 100, 3))), variable.name = "model_type", 
  value.name = "% Positive Predictions") %>% 
  kable() %>% 
  kable_styling(full_width = F)
```
