---
title: "Comparing Logistic Models"
output:
  html_notebook:
    code_folding: hide
    df_print: paged
    highlight: haddock
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '5'
---

```{r setup}
library(tidyverse)
library(ggplot2)
library(caret)
library(modelr)
library(devtools)
load_all("~/savvyR")
```

```{r preprocessing}
load(file = '~/bkt_inv_post_acute/glm_data_preprocessed.RData')
source('~/plot_glm_coef.R')

#----- model fitting stpes -----#
# drop drgs with less than 0.2% non-zero
drg_props<- 
  response_drg  %>%
  select(starts_with("drg")) %>%
  summarise(across(.fns = mean)) %>%
  unlist() 

top_drgs <- which(drg_props > 0.002) %>% names()

# subset df to those drgs
final_df <-
  response_drg %>%
  mutate(age_group = cut(age_at_admit, c(0, 65, 70, 75, 80, 85, 120))) %>%
  select(
    savvy_pid, readmit_30_flag, age_group, gender, ip_days, 
    one_of(top_drgs)
  ) %>% mutate(
    ip_bin_1plus = ifelse(ip_days > 1, 1,0),
    ip_bin_3plus = ifelse(ip_days > 3, 1,0),
    ip_bin_5plus = ifelse(ip_days > 5, 1,0),
    ip_bin_7plus = ifelse(ip_days > 7, 1,0),
    ip_bin_14plus = ifelse(ip_days > 14, 1,0)
  )

final_df <- within(final_df, age_group <- relevel(age_group, "(65,70]"))

# split data into train and test
set.seed(1085)
trainIndex <- createDataPartition(final_df$readmit_30_flag, p=0.8, 
                                  list = FALSE, times = 1)
train_df <- final_df[trainIndex,]
test_df <- final_df[-trainIndex,]
```

## Inspecting Length of Admissions

```{r}
summary(final_df$ip_days)

cat('Total number of members:', n_distinct(final_df$savvy_pid),
'\nNumber of members with 14 days or less length of stay:', length(which(final_df$ip_days <= 14)),
'\nNumber of members with more than 14 days length of stay:', length(which(final_df$ip_days > 14)))

cat('TRAIN DATA:',
'\nTotal number of members:', n_distinct(train_df$savvy_pid),
'\nNumber of members with 14 days or less length of stay:', length(which(train_df$ip_days <= 14)),
'\nNumber of members with more than 14 days length of stay:', length(which(train_df$ip_days > 14)))

cat('TEST DATA:',
'\nTotal number of members:', n_distinct(test_df$savvy_pid),
'\nNumber of members with 14 days or less length of stay:', length(which(test_df$ip_days <= 14)),
'\nNumber of members with more than 14 days length of stay:', length(which(test_df$ip_days > 14)))

final_df %>% 
  filter(ip_days <=30) %>%
  ggplot(aes(x = ip_days)) +
  geom_bar() +
  labs(subtitle = "Length of Inpatient Stay")
```


## Fitting Logistic Models

  * With IP days
  * Without IP days
  * Binned IP days (>1 day, >3 days, >5days, >7 days, >14 days)
  

```{r, echo = T}
# ----- fit the models ----- #
## with IP days
glm_fit <- glm(
  readmit_30_flag ~ . - savvy_pid -ip_bin_1plus -ip_bin_3plus -ip_bin_5plus -ip_bin_7plus -ip_bin_14plus,
  data = train_df,
  family = binomial(link = "logit")
) 

## without IP days
glm_fit_no_IP <- glm(
  readmit_30_flag ~ . -savvy_pid -ip_days -ip_bin_1plus -ip_bin_3plus -ip_bin_5plus -ip_bin_7plus -ip_bin_14plus,
  data = train_df,
  family = binomial(link = "logit")
)

## with binned IP days
glm_fit_bin_IP <- glm(
  readmit_30_flag ~ . -savvy_pid -ip_days,
  data = train_df,
  family = binomial(link = "logit")
)
```

```{r}
## calibration plot function
calib_plot <- function(newdata, fitted.mod, n_buckets = 20){
  
ids_w_preds <- 
  newdata %>%
  add_predictions(model = fitted.mod, type = "response") %>%
  mutate(pred_quant = ntile(pred, n_buckets)) %>%
  select(savvy_pid, pred, readmit_30_flag, pred_quant)

# Data for calibration plot
calib_info <- 
  ids_w_preds %>% 
  group_by(pred_quant) %>%
  summarise(
    true_probs = mean(readmit_30_flag), 
    estim_probs = mean(pred),
  ) %>%
  ungroup() %>%
  rename(ntile = pred_quant)

# Scatter Plot
plot <- calib_info %>%
ggplot(aes(x = estim_probs, y = true_probs)) + 
  geom_hline(yintercept = mean(ids_w_preds$readmit_30_flag), 
             linetype = 2, 
             col = 'red', 
             size = 2) +
  geom_point(size = 3) + 
  labs(
    x = "Estimated Readmission Rate", 
    y = "True Readmission Rate", 
    title = "Model Performace (each point represents 5% of population)"
  ) + 
  geom_smooth(method = "lm") +
  geom_abline(slope = 1, intercept = 0) +
  geom_label(aes(x = 0.025, y = 0.125, label = "Overall Readmission Rate")) +
  xlim(0, 0.24) +
  ylim(0, 0.24) +
  theme_rnd()

  return(list(calib_info, plot))

}

```

## Calibration plots {.tabset .tabset-fade}

#### **True admission rates for the top 5%**
```{r}
rates <- rbind(calib_plot(test_df, glm_fit)[[1]][20,],
  calib_plot(test_df, glm_fit_no_IP)[[1]][20,]) %>%
  rbind(calib_plot(test_df, glm_fit_bin_IP)[[1]][20,]) %>% 
  select(-ntile)

data.frame(models = c('With IP days',
                      'Without IP days',
                      'With IP flags'),
           rates)
```

####

### with IP days
```{r, fig.width=10, fig.height=6}
calib_plot(test_df, glm_fit)[[2]]
```

### without IP days
```{r, fig.width=10, fig.height=6}
calib_plot(test_df, glm_fit_no_IP)[[2]]
```

### binned IP days
```{r, fig.width=10, fig.height=6}
calib_plot(test_df, glm_fit_bin_IP)[[2]]
```


## Coefficient plots {.tabset .tabset-fade}

### with IP days
```{r, fig.width = 10, fig.height = 7}
plot_glm_coef_labs(glm_fit, cutoff = 'positive')
plot_glm_coef_labs(glm_fit, cutoff = 'negative')
```

### without IP days
```{r, fig.width = 10, fig.height = 7}
plot_glm_coef_labs(glm_fit_no_IP, cutoff = 'positive')
plot_glm_coef_labs(glm_fit_no_IP, cutoff = 'negative')
```

### binned IP days
```{r, fig.width = 10, fig.height = 7}
plot_glm_coef_labs(glm_fit_bin_IP, cutoff = 'positive')
plot_glm_coef_labs(glm_fit_bin_IP, cutoff = 'negative')
```


## Cost explorations {.tabset .tabset-fade}
```{r}
library(bigrquery)
pid <- 'research-01-217611'
DB <- 'foc_interoperability.ini_932'
options(httr_oauth_cache = "~/.httr-oauth", httr_oob_default = TRUE)
options(scipen = 20)


test_ids <- data.frame( savvy_pid = test_df$savvy_pid)

bq_project_query(x = pid, query = glue("drop table {DB}_temp_test_ids"))

# upload predicitons to GBQ
ids_tab <- bq_table(
  project = pid,
  dataset = "foc_interoperability", 
  table = "ini_932_temp_test_ids"
)

bq_table_upload(
  x = ids_tab, 
  values = test_ids, 
  create_disposition = "CREATE_IF_NEEDED", 
  fields = as_bq_fields(test_ids)
)
```

```{r}
# Cost/spend information by category for members in the test set
q_preds_cost <-
  glue("
    SELECT
      a.savvy_pid,
      b.spend_30, 
      b.spend_60,
      b.spend_90,
      b.srvc_typ_cd,
      b.hce_srvc_typ_desc
    FROM
      {DB}_temp_test_ids AS a
    LEFT JOIN
      {DB}_costs AS b
    ON 
      a.savvy_pid = b.savvy_pid
  ") 

spend_information <- 
  glue("
    SELECT
      savvy_pid,
      srvc_typ_cd,
      hce_srvc_typ_desc,
      SUM(spend_30) AS spend_30
    FROM
      ({q_preds_cost})
    GROUP BY
      savvy_pid,
      srvc_typ_cd,
      hce_srvc_typ_desc
  ") %>%
  load_bq(db = pid, query = .)
```

```{r}

plot_spend_barplot <- function(newdata, fitted.model, n_buckets = 20){

## add predictions
ids_w_preds <- 
  newdata %>%
  add_predictions(model = fitted.model, type = "response") %>%
  mutate(pred_quant = ntile(pred, n_buckets)) %>%
  select(savvy_pid, pred, readmit_30_flag, pred_quant)

counts_by_group <- 
  ids_w_preds %>%
  mutate(ntile = ifelse(pred_quant <= 18, '<19', pred_quant)) %>% 
  mutate(
    ntile = factor(
      ntile, 
      levels = c("<19", "19", "20"), 
      labels = c("Bottom 90%", "90-95%", "Top 5%")
    )
  ) %>% 
  group_by(ntile) %>%
  summarise(member_count = n())

# Separate 90-95th percentiles and 95-100 percentiles of predictions
spend_w_count <- 
  spend_information %>% left_join(ids_w_preds) %>% 
  mutate(ntile = ifelse(pred_quant <= 18, '<19', pred_quant)) %>% 
  mutate(
    ntile = factor(
      ntile, 
      levels = c("<19", "19", "20"), 
      labels = c("Bottom 90%", "90-95%", "Top 5%")
    )
  ) %>% 
  left_join(
    counts_by_group, 
    by = "ntile"
  )
# costs bar plot
# Plot ip spend in 30 days post-discharge by category 
ip_spend <-
  spend_w_count %>%
  filter(srvc_typ_cd == 'ip') %>%
  group_by(ntile, hce_srvc_typ_desc, member_count) %>%
  summarise(
    total_spend = sum(spend_30)
  ) %>%
  mutate(pmpm = total_spend / member_count)

ip_spend %>% 
  ggplot(
    aes(
      x = fct_reorder(hce_srvc_typ_desc, pmpm), 
      y = pmpm, 
      fill = factor(ntile)
    )
  ) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() + 
  theme_rnd() + 
  scale_fill_rnd() +
  labs(
    x = "Service Description", 
    y = "Spend ($)",
    title = "Average Inpatient Spend by Category", 
    subtitle = "Thirty Days After Discharge", 
    fill = "Risk Group"
  )
}
```

### with IP days
```{r, fig.width = 10, fig.height = 6}
plot_spend_barplot(test_df, glm_fit)
```

### without IP days
```{r,fig.width = 10, fig.height = 6}
plot_spend_barplot(test_df, glm_fit_no_IP)
```

### binned IP days
```{r, fig.width = 10, fig.height = 6}
plot_spend_barplot(test_df, glm_fit_bin_IP)
```



