---
title: "Number of Admissions in 2018 (Negative Binomial)"
output: 
  html_notebook:
    toc: true
    toc_depth: 5
    toc_float: 
      collapsed: false
      smooth_scroll: false
    code_folding: hide
    df_print: paged
    highlight: haddock
    theme: flatly
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = T, warning = F, message = F, comment = NA, fig.align = 'center')

library(MASS)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(caret)
devtools::load_all("~/savvyR")
# install.packages("countreg", repos = 'http://R-Forge.R-project.org')
library(countreg)
library(pscl)
library(kableExtra)
library(stringr)
library(scales)

source("../plot_glm_coef.R")
source("mod_metrics.R")
```

```{r}
# load fitted models 
# models saved due to long runtimes of zinb models
## Negative binomial models
load('~/bkt_inv_post_acute/fit_nb_mods.Rdata')
## Zero Inflated negative binomial models
load('~/bkt_inv_post_acute/fit_zinb_mods.Rdata')
```


## Loading data and pre-processing 

```{r}
# load admission count data
pid <- 'research-01-217611'
DB <- 'foc_interoperability.ini_1085'
options(httr_oauth_cache = "~/.httr-oauth", httr_oob_default = TRUE)
options(scipen = 20)

admit_cnt_df <- load_bq(db = pid,
              query = glue("SELECT *
                            FROM (
                              SELECT *
                              FROM
                                `{DB}_admission_counts_features_2017_2018`
                              WHERE
                                age_2017 >= 65) AS a
                            LEFT JOIN (
                              SELECT *
                              FROM
                                `{DB}_frail_feats_by_quarter`) AS b
                            USING
                              (savvy_pid)"))


## convert  NAs to 0, all on quarterly frailty features
admit_cnt_df[is.na(admit_cnt_df)] <- 0
# save data in bucket
# save(admit_cnt_df, file = '~/bkt_inv_post_acute/admit_cnt_df.rds')
load('~/bkt_inv_post_acute/admit_cnt_df.rds')

```

```{r}
#fix values 
admit_cnt_df$gender <- factor(admit_cnt_df$gender)
admit_cnt_df$age_cat <- cut(admit_cnt_df$age_2017, 
                            breaks = c(65,70,75,80, max(admit_cnt_df$age_2017)+1),
                            labels = c("65-69", "70-74", "75-79", "80+"),
                            right = F)

#filter out members with more than 200 ip days
admit_cnt_df <- admit_cnt_df %>% 
  filter(hosp_days <= 200 & snf_days <= 200)

# split data into train and test
set.seed(1085)
trainIndex <- createDataPartition(1:nrow(admit_cnt_df), p=0.8, list = FALSE)

train_df <- admit_cnt_df[trainIndex,]
test_df <- admit_cnt_df[-trainIndex,]

# center ip days
train_df <- train_df %>% mutate_at(c('hosp_days', 'snf_days'), ~(scale(., scale = FALSE) %>% as.vector))
test_df <- test_df %>% mutate_at(c('hosp_days', 'snf_days'), ~(scale(., scale = FALSE) %>% as.vector))

# save(train_df, test_df, file = '~/bkt_inv_post_acute/nb_train_test_df.Rdata')
load('~/bkt_inv_post_acute/nb_train_test_df.Rdata')


```


## Response Summary
```{r}
# distribution of the response
ggplot(admit_cnt_df, aes(x = admit_counts)) +
  geom_bar() +
  theme_bw() +
  scale_y_continuous(label = comma) +
  labs(
    x = 'Number of admissions in 2018',
    y = 'Member count'
  ) +
  theme(axis.title = element_text(face = 'bold'))

# table of proportions for all the datasets
prop_table <- list()

for(i in 1:3){
  data_list <- list(admit_cnt_df, train_df, test_df)
  table_names <- list('all data', 'train data', 'test data')
  
 temp <- data_list[[i]] %>% 
  mutate(data = table_names[[i]],
         with_admit = ifelse(admit_counts > 0, 1,0)) %>% 
  group_by(data, with_admit) %>% 
  summarize(member_cnt = n(),
            prop = round(100*member_cnt/nrow(data_list[[i]]), 3)) 
 
 prop_table[[i]] <- temp
} 
do.call('rbind',prop_table) %>% 
kable(format = 'html', escape = FALSE) %>% 
  kable_styling() 


# check for response dispersion
c(mean = mean(train_df$admit_counts), 
  var = var(train_df$admit_counts),
  ratio = var(train_df$admit_counts)/mean(train_df$admit_counts))
```


```{r}
# creating feature sets
feat_sets <- list(
  demogs = c('gender', 'age_cat'),
  ip_days = c('hosp_days', 'snf_days'),
  cc = grep('cc_', colnames(train_df), value = T),
  frailty =  grep('frail_.*[[:alpha:]]$', colnames(train_df), value = T),
  frailty_qtr = grep("_Q", colnames(train_df), value = T)
)

# formulas
formulas <- list(
  fm_demog_days = reformulate(termlabels = c(feat_sets$demogs, feat_sets$ip_days),
                              response = 'admit_counts'),
  fm_demog_days_cc = reformulate(termlabels = c(feat_sets$demogs, feat_sets$ip_days, feat_sets$cc),
                              response = 'admit_counts'),
  fm_all = reformulate(termlabels = c(feat_sets$demogs, feat_sets$ip_days, feat_sets$cc, feat_sets$frailty),
                              response = 'admit_counts'),
  fm_all_qtr = reformulate(termlabels = c(feat_sets$demogs, feat_sets$ip_days, feat_sets$cc, feat_sets$frailty_qtr),
                              response = 'admit_counts')
)
```

## Model Fitting  

### Negative Binomial Models {.tabset .tabset-fade}

#### demogs and ip days
```{r}
# demogs and ip days
nb_mod1 <- glm.nb(formulas$fm_demog_days, 
                  data =  train_df)

summary(nb_mod1) 
```

#### demogs, ip days and cc
```{r}
# demographics, ip days, and chronic condition model
nb_mod2 <- glm.nb(formulas$fm_demog_days_cc, 
                  data =  train_df)

summary(nb_mod2) 
```

#### with frailty
```{r}
# demogs and frailty features
nb_mod3 <- glm.nb(formulas$fm_all, 
                  data = train_df)

summary(nb_mod3) 
```

#### with quarterly frailty
```{r}
nb_mod4 <- glm.nb(formulas$fm_all_qtr, 
                  data = train_df)

summary(nb_mod4) 

nb_mods <- list(nb_mod1, nb_mod2, nb_mod3, nb_mod4)
#save(nb_mod1, nb_mod2, nb_mod3, nb_mod4, file = '~/bkt_inv_post_acute/fit_nb_mods.Rdata')
```


### Coefficient plots {.tabset .tabset-fade}
#### demogs and ip days
```{r}
plot_glm_coef(nb_mod1) +
  labs(y = 'Change in Expected Counts')
```

#### demogs, ip days and cc

```{r, fig.height=6, fig.width=9}
plot_glm_coef(nb_mod2) +
  labs(y = 'Change in Expected Counts')
```

#### with frailty 
```{r, fig.height=6, fig.width=9}
 plot_glm_coef(nb_mod3) +
  labs(y = 'Change in Expected Counts')
```

#### with quarterly frailty 
```{r, fig.height= 7, fig.width=9}
plot_glm_coef(nb_mod4, cutoff = 'positive') +
  labs(y = 'Change in Expected Counts')
```
```{r}
plot_glm_coef(nb_mod4, cutoff = 'negative') +
  labs(y = 'Change in Expected Counts')
```


### Rootograms{.tabset .tabset-fade}
#### demogs and ip days

```{r}
rootogram(nb_mod1)
```

#### demogs, ip days and cc

```{r}
rootogram(nb_mod2)
```

#### with frailty
```{r}
rootogram(nb_mod3)
```

#### with quarterly frailty
```{r}
rootogram(nb_mod4)
```

### Zero-inflated Negative Binomial {.tabset .tabset-fade}
Use zero-inflated negative binomial model to account for the presense of excess zeros in the data

#### demogs and ip days
```{r}
# demogs and ip days
## saving zero inflated models due to long runtimes (20mins~2hrs)
zinb_mod1 <- countreg::zeroinfl(formulas$fm_demog_days,
                      data = train_df,
                      dist = "negbin",
                      link = "logit")

summary(zinb_mod1) 
```

#### demogs, ip days and cc
```{r}
zinb_mod2 <- countreg::zeroinfl(formulas$fm_demog_days_cc,
                      data = train_df,
                      dist = "negbin",
                      link = "logit")

summary(zinb_mod2)
```

#### with frailty
```{r}
zinb_mod3 <- countreg::zeroinfl(formulas$fm_all,
                      data = train_df,
                      dist = "negbin",
                      link = "logit")

summary(zinb_mod3) 
```

#### with quarterly frailty
```{r}
zinb_mod4 <- countreg::zeroinfl(formulas$fm_all_qtr,
                      data = train_df,
                      dist = "negbin",
                      link = "logit")

summary(zinb_mod4) 

zinb_mods <- list(zinb_mod1, zinb_mod2, zinb_mod3, zinb_mod4)
# save(zinb_mod1, zinb_mod2, zinb_mod3, zinb_mod4, file = '~/bkt_inv_post_acute/fit_zinb_mods.Rdata')
```


### Coefficient plots {.tabset .tabset-fade}

#### demogs and ip days
```{r}
plot_zinb_coef(zinb_mod1)[[1]] ;plot_zinb_coef(zinb_mod1)[[2]] 
```

#### demogs, ip days and cc

```{r}
plot_zinb_coef(zinb_mod2)[[1]] ;plot_zinb_coef(zinb_mod2)[[2]] 
```

#### with frailty 
```{r, fig.height=6, fig.width=9}
plot_zinb_coef(zinb_mod3)[[1]] ;plot_zinb_coef(zinb_mod3)[[2]] 
```

#### with quarterly frailty 
```{r, fig.height=7, fig.width=9, dpi = 750}
plot_zinb_coef(zinb_mod4)[[1]] ;plot_zinb_coef(zinb_mod4)[[2]] 
```


### Rootograms {.tabset .tabset-fade}

#### demogs and ip days
```{r}
rootogram(zinb_mod1)
```

#### demogs, ip days and cc
```{r}
rootogram(zinb_mod2)
```


#### with frailty
```{r}
rootogram(zinb_mod3)
```

#### with quarterly frailty
```{r}
rootogram(zinb_mod4)
```

## Model Comparison

### Metrics on test data {.tabset .tabset-fade}

#### admit_counts > 0
```{r}
metric_smry(nb_mods, zinb_mods, count = 0, formulas, cutoff = 0.5)
```

#### admit_counts > 1
```{r}
metric_smry(nb_mods, zinb_mods, 1, formulas, cutoff = 0.5)
```

#### admit_counts > 2
```{r}
metric_smry(nb_mods, zinb_mods, 2, formulas, cutoff = 0.5)
```


### % Positive Predictions {.tabset .tabset-fade}

#### admit_counts > 0

```{r}
pospred_smry(nb_mods, zinb_mods, count = 0, cutoff = 0.5)
```


#### admit_counts > 1

```{r}
pospred_smry(nb_mods, zinb_mods, count = 1, cutoff = 0.5)
```

#### admit_counts > 2

```{r}
pospred_smry(nb_mods, zinb_mods, count = 2, cutoff = 0.5)
```






## Exploring zinb_mod4

Model with gender, age, ip days, chronic conditions, frailty categories by quarter as predictors.
Lowest RMSE and PPV

```{r}
mod_viz <- mod_sim_viz(test_df, zinb_mod4); mod_viz[[3]] 
```


```{r, include= FALSE}
  
pred <- predict(zinb_mod4, test_df, type = "response")

# generate simulated negbinom variates
sim_pred <- rnbinom(n = test_df$admit_counts, mu = pred, size = zinb_mod4$theta) 

# creating obs vs simulated df
obs <- table(test_df$admit_counts) %>% data.frame() %>% mutate(type = 'actual')
sim <- table(sim_pred) %>% data.frame() %>% mutate(type = 'simulated')
names(sim) <- c("Var1", "Freq", "type")

tmp <- as.data.frame(rbind(obs, sim)) %>% mutate(with_admit = ifelse(Var1 != 0, 1,0)) 

tmp %>% 
  group_by(type, with_admit) %>% 
  summarise(count = sum(Freq)) %>% 
  group_by(type) %>% 
  mutate(prop = round(count/sum(count)*100,3)) %>% 
kable(format = 'html', escape = FALSE) %>% 
  kable_styling() 
```


```{r}
# create df with actual and simulated counts
plot_df <- data.frame(pred = sim_pred, actual = test_df$admit_counts)

# actual vs. simulated data on test data
mod_result <- plot_df %>%
  mutate(abs_diff = abs(actual - pred),
         cnt_grp = case_when(
           actual == 0 ~ '0',
           actual == 1 ~ '1',
           actual == 2 ~ '2',
           actual >= 3 ~ '3+'
         )) 

mod_acc_smry <- mod_result %>% 
  group_by(cnt_grp) %>% 
  dplyr::summarise(mem_cnt = n(),
            proportion = round(n()/nrow(mod_result)*100,3),
            correct_pred_mem_cnt = length(which(abs_diff == 0)),
            pred_exact_admit = round(sum(ifelse(abs_diff == 0, 1,0))/mem_cnt,4),
            pred_within_1_admit = round(sum(ifelse(abs_diff <= 1, 1,0))/mem_cnt,4),
            pred_within_2_admit = round(sum(ifelse(abs_diff <= 2, 1,0))/mem_cnt,4),
            pred_within_3_admit = round(sum(ifelse(abs_diff <= 3, 1,0))/mem_cnt,4)
            ) 
  
mod_acc_smry %>% 
kable(format = 'html', escape = FALSE) %>% 
  kable_styling() 

library(ggrepel)

mod_acc_smry[,-c(2:4)] %>% gather(pred_acc, values, -cnt_grp) %>% 
  ggplot(aes(y = values, x = pred_acc, col = cnt_grp)) +
  geom_line(aes(group = cnt_grp)) +
  geom_point(shape = 15, size =2) +
  geom_label_repel(aes(label = scales::percent(values, .11)), size = 3)+
  scale_y_continuous(labels = scales::percent)+
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.5)
  )+
  labs(y = 'Proportion of Predictions',
       x = '',
       title = 'Model prediction on test data',
       subtitle = 'Zero-inflated Negative Binomial (demogs, ip days, chronic conditions, frailty by quarter)',
       col = 'Number of\nAdmissions')
  
```




