---
title: "Number of Admissions in 2019 (non elective admissions, negative Binomial)"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    highlight: haddock
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(data.table)
library(tidyverse)
library(glue)
library(drake)
library(devtools)
library(here)
library(caret)
devtools::load_all("~/savvyR")
devtools::load_all("~/ini_620_home_complex_risk_strat/homeriskstratR")
cache <- new_cache(path = "/storage_pool/cache/")

options(httr_oauth_cache = "~/.httr-oauth", httr_oob_default = TRUE)
options(scipen = 20)
```


```{r load data}
load('~/bkt_inv_post_acute/metrics_2018.Rdata')
load('~/bkt_inv_post_acute/metrics_2018_lag.Rdata')
load('~/bkt_inv_post_acute/data_2018.Rdata')
loadd(models_w_ckd_stages, cache = cache)

start_dt <- '2018-01-01'
end_dt <- '2018-12-31'
pid <- 'research-01-217611'
DB <-  "foc_interoperability.ini_1085"  
```

```{r functions, echo=FALSE}
source(here::here("code/modeling/admission_negbinom_model", "mod_metrics.R"))

metrics_table <- function(data, perc){
data %>% 
  filter(target_perc == perc) %>%
  select(-target_perc) %>% 
  mutate_at(vars(-c(1:5)), ~round(.,2)) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c('striped', 'hover', 'condensed')) %>% 
  collapse_rows(1, valign = 'middle') 
}

make_metrics_table = function(admit_cnt_df, pids){
  
  df <- admit_cnt_df %>% filter(savvy_pid %in% pids) %>% 
    dplyr::select(savvy_pid, admit_counts,snf_admits,
                  total_spend, hosp_spend, snf_spend, mm_2019, date_of_death) %>% 
    filter(mm_2019 > 0) %>% 
    mutate(death_flag = ifelse(is.na(date_of_death)| date_of_death > '2020-01-01', 0, 1)) 
  
  # compute metrics for the predicted members 
  list(df %>% 
    summarise(PMPM_total = sum(total_spend, na.rm = T)/(sum(mm_2019)),
              PMPM_hosp = sum(hosp_spend, na.rm = T)/(sum(mm_2019)),
              PMPM_snf = sum(snf_spend, na.rm = T)/(sum(mm_2019)),
              admits_1k = (sum(admit_counts,na.rm = T)/(n_distinct(savvy_pid)))*1000,
              snf_admits_1k = (sum(snf_admits,na.rm = T)/(n_distinct(savvy_pid)))*1000,
              mortality_rate = sum(death_flag)/(n_distinct(savvy_pid))))
  
}

# function to get columns for metrics of identified frail population

string_serious <- select(data_2018, contains('serious')) %>% names() %>% toString()

get_metric_vars_query <- function(DB, pop_query, tbl_name, year_nbr, with_lag = F){
  serious_year <- if(with_lag == F){
    year_nbr
  } else {noquote(paste0(year_nbr,10))}
  
  glue("SELECT savvy_pid, admit_counts, mm_{year_nbr + 1}, total_spend, 
               hosp_spend, snf_spend, snf_admits, date_of_death, 
               age_{year_nbr}, gender,
               frail_Falls_and_Fractures,
               frail_Pressure_Ulcers_and_Weight_Loss, frail_Incontinence, frail_Anxiety_and_Depression,
               frail_Mobility_Problems, frail_Dependence_and_Care,frail_Dementia_and_Delirium, {noquote(string_serious)}
        FROM 
        (SELECT * 
        FROM `{DB}_{tbl_name}`
        WHERE savvy_pid IN (SELECT savvy_pid FROM ({pop_query})) 
        AND age_{year_nbr} >= 65) as a
        
        LEFT JOIN (
        SELECT savvy_pid, {noquote(string_serious)}
        from {DB}_serious_conditions_{year_nbr}
        ) as b
        
        USING (savvy_pid)
      ")
}

add_race <- function(q_ids, pop_df){
  
  ## q_ids = table or query with savvy_pid
  ## pop_df = demographic table of identified population
  
  race <- load_bq(pid, query = race_cms(q_ids))

  cols <- colnames(race)[endsWith(colnames(race), '_flag') == TRUE]
  
  race <- race %>% dplyr::select(savvy_pid,cols) %>% 
    tidyr::gather(race, flag, -savvy_pid) %>% 
    filter(flag == 1) %>% 
    full_join(data.frame(savvy_pid = pop_df$savvy_pid))

  race$race[is.na(race$race)] <- 'no_data'

  ## duplicates with inconsistent races
  dups <-  race %>% 
    group_by(savvy_pid) %>% 
    filter(n()>1) %>% 
    arrange(savvy_pid) %>% 
    dplyr::select(savvy_pid) %>% 
    distinct()
  
  inconsistent_race <- data_frame(savvy_pid = as.vector(dups$savvy_pid),
    race_dup = 'inconsistent_data')
  
  race_df = full_join(race, inconsistent_race, by = "savvy_pid") %>% 
    mutate(race_final = ifelse(is.na(race_dup), race, race_dup)) %>% 
    dplyr::select(savvy_pid, race_final) %>% distinct()
  
  race_df$race_final <- gsub("_flag", "", race_df$race_final)
  
  final_data <- left_join(pop_df, race_df)
}

# preprocessing function for admit count dataframes
preprocess_data = function(df)
{
  # create hcc_over5_flag
  df <- df %>% 
    mutate(hcc_over5_flag = ifelse(hcc_count > 5, 1,0),
           prior_admit_1plus = ifelse(admit_6mos_prior_counts >= 1, 1, 0),
           prior_admit_2plus = ifelse(admit_6mos_prior_counts >= 2, 1, 0))
  
  # vars not included
  vars_not_included <- c("hosp_spend", "snf_spend", "total_spend", "snf_admits",
                         "date_of_death", "mm_2019", "gender", "hcc_count")
  df <- df %>% 
    mutate_if(!(names(.) %in% vars_not_included), ~ifelse(is.na(.), 0, .))
  
  df$gender <- factor(df$gender)
  df$age_cat <- cut(
    df$age_2018,
    breaks = c(65, 70, 75, 80, max(df$age_2018) +
                 1),
    labels = c("65-69", "70-74", "75-79", "80+"),
    right = F
  )
  dmy <- dummyVars(~ age_cat, data = df)
  dmy_df <- predict(dmy, newdata = df)
  colnames(dmy_df) <-  make.names(colnames(dmy_df))
  
  df <- cbind(df, dmy_df)
  
  # use LACE index LOS cutoffs as breaks
  df$hosp_cat <- cut(
    df$hosp_days,
    breaks = c(0, 1, 2, 3, 4,7, 14, max(df$hosp_days) +
                 1),
    labels = c("0", "1", "2", "3", "4-6", "7-13", ">=14"),
    right = F
  )
  
  
  # breaks based on medicare coverage days
  df$snf_cat <- cut(
    df$snf_days,
    breaks = c(0, 21, 101, max(df$snf_days) +
                 1),
    labels = c("0-20", "21-100", ">100"),
    right = F
  )
  
  #filter out members with more than 200 ip days
  # df <- df %>% 
  #   filter(hosp_days <= 200 & snf_days <= 200)
  return(df)

}
```

# {.tabset .tabset-fade}
## Data from 2018 - 2019
### Using % positive predicted dependent threshold {.tabset .tabset-fade}

#### 5% positive predicted
```{r}
metrics_table(metrics_uniform_perc18, 0.05)
```


#### 2% positive predicted
```{r}
metrics_table(metrics_uniform_perc18, 0.02)
```


### Threshold from cutoff > 0 (% positive predicted dependent) {.tabset .tabset-fade}

#### 5 % positive predicted
```{r}
metrics_table(metrics_fixed_0_18, 0.05)
```

#### 2% positive predicted
```{r}
metrics_table(metrics_fixed_0_18, 0.02)
```

## Data from 2018, simulated lag, data up to Oct. 2018
### Using % positive predicted dependent threshold {.tabset .tabset-fade}

#### 5% positive predicted
```{r}
metrics_table(metrics_lag_uniform_perc18, 0.05)
```


#### 2% positive predicted
```{r}
metrics_table(metrics_lag_uniform_perc18, 0.02)
```


### Threshold from cutoff > 0 (% positive predicted dependent) {.tabset .tabset-fade}

#### 5 % positive predicted
```{r}
metrics_table(metrics_lag_fixed_0_18, 0.05)
```

#### 2% positive predicted
```{r}
metrics_table(metrics_lag_fixed_0_18, 0.02)
```

## Other Algorithms

```{r, eval = FALSE}
# MA population 2018 ------------------------------------------------------
# continuously enrolled members in 2018
cohort_ma <- glue("
   SELECT
     distinct savvy_pid
   FROM
     foc_interoperability.ini_1085_pop_ce_2018
               
")

# pop_2018 <- load_bq(pid, cohort_ma) 



######## NaviHealth population
# first part of table name
db_ma <- 'foc_interoperability.ini_1085_ma_2018'

claims_ma <- miniov_medical(start_dt, end_dt, cohort_ma)

hcc_table_ma <- glue("{db_ma}_cms_raf")

## frail, medically complex cohort using NaviHealth criteria
q_navi_ma <- 
  navi_frail_pop(
    hcc_table_ma,
    claims_ma
  )

# load navihealth population
pop_navi_2018 <- load_bq(pid, 
                         query = get_metric_vars_query(DB, q_navi_ma, year_nbr = 2018,
                                                       tbl_name = "admission_counts_features_non_elect_2018_2019"))

pop_navi_2018 <- add_race(q_navi_ma, pop_navi_2018) %>% distinct()

```

```{r}
######## 620 population
ini_620_population_codes <- read_csv("../../../data/ini_620_population_codes.csv")

## get 2018 cohort
q_data_620 <- get_cohort(start_dt = '2018-01-01', ini_620_codes = ini_620_population_codes)

data_620 <- load_bq(pid, q_data_620)

# required columns for identified 620 population
pop_620_2018 <- load_bq(pid, 
                        query = get_metric_vars_query(DB, q_data_620,  year_nbr = 2018,
                                                      tbl_name = "admission_counts_features_non_elect_2018_2019"))


pop_620_2018 <- add_race(q_data_620, pop_620_2018)
```

```{r}
query_2018 <-  glue("SELECT *
              FROM 
              (
                SELECT *
                FROM
                  `{DB}_admission_counts_features_non_elect_2018_2019`
              ) AS a
              LEFT JOIN (
                SELECT *
                FROM
                  `{DB}_frail_feats_by_quarter_2018`) AS b
              USING
                (savvy_pid)
              LEFT JOIN (
                SELECT * EXCEPT (total_raf)
                FROM
                  `{DB}_serious_conditions_2018`) as c
              USING (savvy_pid)")

# data_2018 <- load_bq(pid, query_2018)
names(data_2018)[colnames(data_2018) == 'CKD_other'] <- 'CKD_Other'

data_2018 <- preprocess_data(data_2018)

pop_mod_ids <- get_threshold(fitted_model = models_w_ckd_stages, train_data = data_2018, 
                  target_perc = 0.02, cutoff = 0)[[2]]$savvy_pid

tbl_name <-  "admission_counts_features_non_elect_2018_2019"

q_mod_ids <- glue("SELECT *
                FROM
                  `{DB}_{tbl_name}`
                WHERE
                  savvy_pid IN ({toString(pop_mod_ids)})")

pop_mod_2018 <- load_bq(pid, 
                        query = get_metric_vars_query(DB, q_mod_ids,
                                                      tbl_name, year_nbr = 2018))

pop_mod_2018 <- add_race(q_mod_ids, pop_mod_2018)

```

```{r}
query_lag <-  glue("SELECT *
              FROM 
              (
                SELECT *
                FROM
                  `{DB}_admission_counts_features_non_elect_2018_trunc`
              ) AS a
              LEFT JOIN (
                SELECT *
                FROM
                  `{DB}_frail_feats_by_quarter_201810`) AS b
              USING
                (savvy_pid)
              LEFT JOIN (
                SELECT * EXCEPT (total_raf)
                FROM
                  `{DB}_serious_conditions_201810`) as c
              USING (savvy_pid)")

data_lag <- load_bq(pid, query_lag)
names(data_lag)[colnames(data_lag) == 'CKD_other'] <- 'CKD_Other'

data_lag <- preprocess_data(data_lag)

pop_lag_ids <- get_threshold(fitted_model = models_w_ckd_stages, train_data = data_lag, 
                  target_perc = 0.02, cutoff = 0)[[2]]$savvy_pid

tbl_name <-  "admission_counts_features_non_elect_2018_trunc"

q_lag_ids <- glue("SELECT *
                FROM
                  `{DB}_{tbl_name}`
                WHERE
                  savvy_pid IN ({toString(pop_lag_ids)})")

pop_lag_2018 <- load_bq(pid, 
                        query = get_metric_vars_query(DB, q_lag_ids,
                                                      tbl_name, year_nbr = 2018))

pop_lag_2018 <- add_race(q_lag_ids, pop_lag_2018)

# save(data_2018,
#      data_lag,
#      file = '~/bkt_inv_post_acute/data_2018.Rdata')

save(pop_navi_2018,
     pop_620_2018,
     pop_mod_2018,
     pop_lag_2018,
     pop_2018,
     file = '~/bkt_inv_post_acute/other_algo_2018.Rdata')
```


### on 2018-2019 data

```{r}
load('~/bkt_inv_post_acute/other_algo_2018.Rdata')

pop_mod_2018 <- pop_mod_2018 %>% mutate(model = 'our model')
pop_lag_2018 <- pop_lag_2018 %>% mutate(model = 'our model (on data with lag)')
pop_navi_2018 <- pop_navi_2018 %>% mutate(model = 'NaviHealth')
pop_620_2018 <- pop_620_2018 %>% mutate(model = 'Home Complex Care')

pop_all <- rbind(pop_mod_2018, pop_lag_2018, pop_navi_2018, pop_620_2018)

models <- unique(pop_all$model)

x <- as.data.frame(t(sapply(models, function(x) 
  (filter(pop_all, model == x) %>% 
     make_metrics_table(., .$savvy_pid)) %>% 
    as.data.frame() %>% 
    dplyr::mutate(model = paste(x)) %>% 
    mutate_at(vars(-c(7)), ~round(.,2))
  ))) %>%  stack() %>% unstack()


temp = pop_all %>% 
  group_by(model) %>% 
  summarise(proportion = n()/ length(pop_2018)*100) %>% 
  mutate_at(vars(-c(1)), ~round(.,2)) 

ord <- c(names(x)[-7], 'proportion')

z <- left_join(x,temp) %>% 
  select(model, everything()) %>%
  tidyr::gather(variable, value, -model) %>% 
  spread(model, value) %>% 
  slice(match(ord, variable))


z %>% 
kable() %>% 
kable_styling(bootstrap_options = c('striped', 'hover', 'condensed'))
  
```



```{r}

perc <- function(x){
  paste(round(x*100,2))
}

summary_tab <- pop_all %>% 
  select(-c(savvy_pid, mm_2019, contains('spend'), snf_admits, date_of_death)) %>% 
  mutate(admit_flag = ifelse(admit_counts >0, 1, 0)) %>% 
  group_by(model) %>% 
  tidyr::gather(key, value, -c(age_2018, admit_counts, admit_flag, model)) %>% 
  select(model, key,value) %>% 
  group_by(model, key, value) %>% 
  summarize(count = n()) %>% 
  mutate(prop = round(100*count/sum(count),1),
         total = sum(count)) %>% 
  select(-count) %>% filter(value != 0)

pop_all %>% 
  group_by(model) %>% 
  summarize(age_summary = paste(round(mean(age_2018),1), '\u00B1', round(sd(age_2018),1), ',', median(age_2018)))


```

```{r}
pop_all %>%
  ggplot(aes(x = race_final, group = model, y = ..prop.., fill = model)) +
  geom_bar(position = 'dodge', stat = 'count', aes(y = ..prop..)) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme(
    legend.position = 'bottom',
    axis.text = element_text(angle = 45, hjust = 1)
    )


pop_all2 %>% 
  ggplot(aes(x = model, y = age_2018, fill = model)) +
  geom_boxplot() +
   theme(
    legend.position = 'bottom'
  ) +
  ylim(20,100)
```

